Multi-task learning in Computer Vision

### 2020-02-26 3-Image-classification.tex
Start seminar mode

\chapter{Image classification}
Image classification is one of the essential modern computer vision problems, where the goal is to create a model that can classify an input image into one of a set of pre-defined classes. Before the popularization of applying large Convolutional Neural Networks for this task, the most successful way of solving the problem was to use some algorithm for finding feature descriptors in a set of images to construct a Visual Bag of Words.  A linear classifier, like a Support Vector Machine, would then do the classification using the Bag of Words representation of an image. These days nearly all approaches are based on using deep CNNs, and working CNNs were deployed already in 1998 on character recognition in the form of LeNet \citep{leNet}.

\section{ImageNet}

ImageNet  \citep{imagenet}  is perhaps the most significant dataset for image classification and especially the ImageNet Challenge \citep{ILSVRC}, which is a challenge for a collection of 1000 classes from the ImageNet dataset for image classification using 1.2 million training and 150 thousand test images of the entire ImageNet dataset. In 2012 the winning model, AlexNet \citep{alexNet}, showed that it was possible to train a deep CNNs efficiently using GPUs. Since 2012 all top-performing models showed some new improvements on how to create the most performant network architecture, for example, VGGNet from the year 2014 and ResNet \citep{resNet} from 2015, both of which have been popular models to use for Transfer Learning since.

Human accuracy on the ImageNet challenge is about 5.1\% \citep{imageNet_summary}, and ResNet achieves a top-5 error rate of 3.57\% \citep{resNet} and newer architectures even lower, but this still does not mean that image classification is a solved problem. The human performance experiment found that many of the human errors are caused by not having expert information in, for example, identifying animal species or not even being aware of the existence of a class \citep{imageNet_summary}. ObjectNet \citep{objectNet} is a dataset designed to test image classifiers with a focus on their generalizability. It contains many classes that also exist in the ImageNet dataset. However, they are in unexpected locations or have an unexpected pose, causing the high accuracy image classifiers trained on ImageNet to experience a 40-45\% accuracy drop when evaluating them on the ObjectNet images of classes shared between ImageNet and ObjectNet. This kind of adjustment is relatively easy for a human, and it shows that while the classifiers are good, they are by no means perfect.

@@

Unlike the ImageNet challenge, most real-world tasks do not have such an abundance of data for all possible classes. Still, to achieve the highest accuracies, they require models that are equal in terms of complexity to those that have top accuracies problems on the scale of the ImageNet classification. For this reason, many CNN classifiers, irrespective of the problem, feature one of the ImageNet classifiers in the model architecture. Even though some datasets may contain a large number of images per class, using a pre-trained classifier as a basis often produces a better final classifier by applying fine-tuning \citep{betterTransfer}.

@@

\section{ResNet}
ResNet is one of the first effective and very deep Convolutional Neural Network architecture that was presented in 2015 and won the ImageNet challenge. Prior to the publication of ResNet, the most powerful networks were relatively shallow, like VGGNet, which has only 19 layers. One of the big issues relating to training deep networks is vanishing gradients, where gradients disappear when they are backpropagated through many layers \citep{wideResNet}. ResNets utilize residual connections around bottleneck building blocks, which allow for the networks to contain many more layers than those without them, the largest network presented in the original ResNet paper was 152 layers, totaling for around 8x increase in the number of layers when compared to earlier networks \citep{resNet}.

\begin{figure}[h!] 
\centering 
\includegraphics[width=0.8\textwidth]{imgs/resnet-block.png}
\caption{Resnet building block \citep{resNet}}
\end{figure}

The ResNet block architecture allows for the blocks to learn the identity function more efficiently by trying to learn the residual function instead of the direct mapping. Normally a mapping is learned between ${x}$ and ${y}$ using a function ${H(x)}$ but by the same token, we can learn a residual transformation, ${F(x) = H(x) - x}$, where ${H}$ is the mapping of two or more network layers. Both of these approaches should approximate the same functions; the difference is in how easy it is for the network to learn. Learning a transformation of ${F(x) = 0}$ would intuitively seem easier for a neural network than learning ${F(x) = x}$. As can be seen from the success of the ResNets compared to non-residual networks, this is what allows for creating very deep networks. If the transformation changes the input size, a matrix ${W}$ is necessary to map the input to the same dimensions, generating the final formula for a residual block ${y = F(x, \{W_i\}) + W_s x}$.

Many variants of ResNet exists, such as wide ResNets \citep{wideResNet}, ResNeXt \citep{resNext} and others. Also the DenseNet \citep{denseNet} is heavily inspired by ResNets. ResNet-50, ResNet-34, ResNet-101, and ResNet-152 are still some of the most popular models to use when a pre-trained ImageNet trained backbone is needed for some part of a classifier as they produce good results and do not contain too many parameters compared to some other architectures.

\section{Improving model performance}
Improving model performance is not easy, but using the various types of skip-connections, as the ResNet residual blocks use, it is possible to increase the size of the network to massive sizes. A 557 million parameter model called GPipe \citep{gPipe} takes the model scaling to the extreme and requires some unique parallelism libraries to train the model. It is still only slightly better than previous models, showing that size is not the only thing that matters.

\begin{figure}[h!] 
\centering 
\includegraphics[width=0.8\textwidth]{imgs/scaling-networks.png}
\caption{Various ways of scaling network architectures \citep{efficientNet}}
\end{figure}

There are three main ways to scale up a network, as shown in figure 3.3. Scaling by depth means adding more layers to the model, allowing for more complex dependencies to be captured by the model. For example, ResNet-1000 is a very deep type of ResNet, but it has similar performance to a ResNet-101, so there are diminishing returns when trying to scale up by depth \citep{efficientNet}. Scaling by width means increasing the number of channels in the layers, and it is especially popular when optimizing smaller size models, such as MobileNet \citep{mobileNet}. Wide ResNets \citep{wideResNet} increase the width of the ResNet blocks and allow for better features and easier training.

\section{EfficientNet}
EfficientNet models form a family of models ranging from EfficientNet-0 to EfficientNet-7 that were generated by smartly scaling existing convolutional models to optimize them for efficiency \citep{efficientNet}. The more complex models are created by using compound scaling on the EfficientNet-0 model, where the width, depth, and the resolution of the network are scaled using a factor ${\phi}$. The search of ${\phi}$ is an optimization problem, where the goal is to optimize for both accuracy and number of floating-point operations. This search is only done on the base model because the neural architecture search used to find the parameter becomes very expensive as the model size increases \citep{efficientNet}.

The proposed EfficientNet models use mobile inverted bottleneck (MBConv) blocks used in MobileNetV2 \citep{mobileNetv2} in constructing the base model. The MobileNet uses a Depthwise Separable Convolution, where a depthwise convolution and a pointwise convolution are applied in sequence. The depthwise convolution applies ${d_i}$ ${k x k}$ filters to the input, where ${d_i}$ is the input channels and ${k}$ is the kernel size leading to an output channel count ${d_j = d_i}$. A normal convolutional layer would apply multiple filters having ${M}$ channels with the computational cost of ${h_i w_i d_i d_j k^2}$, whereas the depthwise convolution only has a cost of ${h_i w_i d_i (k^2 + d_j)}$, reducing the cost by ${k^2}$. The result of the depthwise convolution then runs through a pointwise convolution, where a 1 x 1 x ${d_j}$ 1d convolution is applied to get the final output as a linear combination of the channels.

>>>>

\chapter{Transfer Learning}

@@

@@

### 2020-02-26 5-Multitask.tex
Start seminar mode

\chapter{Multi task learning}
Cover some examples of MTL in Computer Vision, differences to Transfer Learning, why it might work, why we might see detrimental results and why it might be beneficial (results/performance)
\section{Definition}
\section{Benefits}
\section{Hard parameter sharing}
\section{Soft parameter sharing}
\section{Special multi-task approaches}
\section{Multi-task learning with little data}
\section{What and when to share}
\section{Succesful multi-task learning applications}

>>>>

\chapter{Multi-task learning}
Maybe drop some sections if it becomes too long.
\section{Definition}
Formally define multi-task learning similar to Transfer Learning using the original multi-task paper \citep{origMultitask}, and a newer view on the domain \citep{surveyOnMultiTask}.
\section{Benefits}
Inference time, regularization, smaller models etc. Weighing different tasks and getting information from the loss functions \citep{lossWeighting} \citep{usingUncertaintyToWeighLosses}
\section{Soft parameter sharing}
Define soft parameter sharing between tasks, show an example from \citep{mutualLearning}.
\section{Hard parameter sharing}
Show how multiple heads can be combined to produce outputs for various tasks using different heads, show results from sharing all/some layers. \citep{visualPerson} \citep{selfDriving} \citep{healthyDrink}
\section{Special multi-task learning techniques}
Show how an attention layer can be added to augment the performance in various tasks in a multi-task setting \citep{multiTaskAttention}. Also cross-stich networks \citep{crossStitch}.
\section{Multi-task learning with little data}
Why it can be beneficial with little data \citep{surveyOnMultiTask} Show some results where it is beneficial on little data.
\section{What and when to share}
Some ways to decide when multi-task learning might be a good idea and how to evaluate which tasks might be good to learn in a multi-task setting \citep{taskonomy} \citep{whichTasks}.
Show how this problem can be similar to deciding whether Transfer Learning is a good idea \citep{whatAndWhereToTransfer} \citep{transferringMidLevelRepresentations}

### 2020-02-26 HY-CS-main.tex
Add transfer learning chapter
Move chapters to their own files, add some model scaling info
Start seminar mode

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% STEP 2:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Set up personal information for the title page and the abstract form.
%% Replace parameters with your information.
\title{Weather classification using Convolutional Neural Networks}

@@

% Set supervisors and examiners, use the titles according to the thesis language
% Prof. 
% Dr. or in Finnish toht. or tri or FT, TkT, Ph.D. or in Swedish... 
\supervisors{Prof. Laura Ruotsalainen}
\examiners{Prof.~D.U.~Mind, Dr.~O.~Why}

@@

\classification{\protect{\ \\
\  General and reference $\rightarrow$ Document types  $\rightarrow$ Surveys and overviews\  \\
\  Applied computing  $\rightarrow$ Document management and text processing  $\rightarrow$ Document management $\rightarrow$ Text editing
}}

@@

\begin{otherlanguage}{english}
\begin{abstract}
Use this otherlanguage environment to write your abstract in another language if needed.

Topics are classified according to the ACM Computing Classification System
(CCS), see 
\url{https://www.acm.org/about-acm/class}:
check command \verb+\classification{}+. A small set of paths (1--3) should be used, starting from any top nodes
referred to bu the root term CCS leading to the leaf nodes. The elements
in the path are separated by right arrow, and emphasis of each element individually can be indicated
by the use of bold face for high importance or italics for intermediate
level. The combination of individual boldface terms may give the reader
additional insight. 
\end{abstract}
\end{otherlanguage}

@@

% BSc instructions
%\include{bsc_finnish_contents}
%\include{bsc_english_contents}
% MSc instructions
%\include{msc_finnish_contents}
\include{msc_english_contents}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% STEP 2:
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Set up personal information for the title page and the abstract form.
%% Replace parameters with your information.
\title{Multi-task learning in Computer Vision}

@@

% Set supervisors and examiners, use the titles according to the thesis language
% Prof. 
% Dr. or in Finnish toht. or tri or FT, TkT, Ph.D. or in Swedish... 
\supervisors{Prof.~D.U.~Mind}
\examiners{Prof.~D.U.~Mind, Dr.~O.~Why}

@@

@@

@@

% BSc instructions
%\include{bsc_finnish_contents}
%\include{bsc_english_contents}
% MSc instructions
%\include{msc_finnish_contents}
\include{1-Introduction}
\include{3-Image-classification}
\include{5-Multitask}
\include{7-Experiments}
\include{8-Conclusions}
% \include{msc_english_contents}
